Very long and verbose title

### 2022-03-28 docs/holodeck-paper/src/ms.tex
manuscript updates

\newcommand{\thardf}{\tau_f}
\newcommand{\holodeck}{\texttt{holodeck}}
\newcommand{\python}{\texttt{python}}
\newcommand{\mmbulge}{{$M_\textrm{BH}$--$M_\textrm{bulge}$}}

@@

    Here, the GW strain from a circular binary is \citep[][Eq.~7]{Sesana+2008} (sky and polarization averaged),
        \begin{equation}
        \label{eq:source_strain}
        h_s = \frac{8}{10^{1/2}} \frac{\left(G\mchirp\right)^{5/3}}{c^4 \, d_L}
            \left(2 \pi f_r \right)^{2/3},
        \end{equation}
    for a luminosity distance $d_L$.

@@

        Define the (comoving-) number density of sources as, $n \equiv dN / dV_c$.

        From \citet[][Eqs.~5/8]{Phinney-2001},
        	\begin{align}
        	h_c^2(f) = \int_0^\infty \!\! dz \; \frac{d^2 N}{dz \, d\ln f_r} \; h_s^2\lr{f_r}.
        	\end{align}
        From \citet[][Eq.~6]{Sesana+2008} we can write,
        	\begin{align}
        	\frac{d^2 N}{dz \, d\ln f_r} = \frac{d n_c}{dz} \frac{dz}{dt} \frac{dt}{d\ln f_r} \frac{d V_c}{dz}.
        	\end{align}
        The binary hardening timescale is given by the expression,
        \begin{equation}
            \thardf \equiv \diffp{t}{{\ln\!f_r}} = \frac{f_r}{\partial f_r / \partial t}
        \end{equation}
        From cosmography \citep[e.g.][]{Hogg-1999},
        	\begin{align}
        	\frac{dz}{dt} = H_0 \lr{1+z} E(z) \\
        	\frac{d V_c}{dz} = 4\pi \frac{c}{H_0} \frac{d_c^2}{E(z)} \\
          d_L = d_c \, (1+z)
        	\end{align}
        Putting this together we have,
        	\begin{align}
        	h_c^2(f) & = \int_0^\infty \!\! dz \; \frac{dn_c}{dz} \, h^2\lr{f_r} \, 4\pi c \, d_c^2 \lr{1+z} \, \thardf.
        	\end{align}
        % If this is being calculated from a finite volume (e.g.~simulation), then we can associate the comoving-number-density of sources at a given redshift with the number of binaries per-unit comoving volume, i.e.,
        % 	\begin{align}
        % 	\frac{dn_c}{dz} \, dz = N_\trt{fv}(z)/V_\trt{c,fv}.
        % 	\end{align}

@@

        Define the (comoving-) number density of sources as, $n \equiv dN / dV_c$, and the distribution function of sources as $f(M,a,z) = d^2 n(M,a,z) / dM da$.  Here $M$ is the total mass of each systems, $a$ is the binary separation, and $z$ is the redshift.  Note that binary frequency $f$ could be used almost interchangeably with $a$, and that both evolve as a function of time or equivalently redshift.  This formalism can easily be extended to include other parameters, such as mass-ratio and eccentricity ($q$ and $e$).  We can write the conservation equation for binaries as of function of redshift as,
        \begin{equation}
            \label{eq:conservation}
            \diffp{f}{z} + \diffp{}{m} \lrs{f \diffp{m}{z}} + \diffp{}{a} \lrs{f \diffp{a}{z}} = S_f(m, a, z).
        \end{equation}
        Here $S_f$ is a source/sink function that can account for the creation or destruction of binaries.  Here we assume that all binary `formation' is encapsulated from binaries moving from one part of parameter space (i.e.~large separations and redshifts) to other parts of parameter space (i.e.~smaller separations and redshifts).  Thus we set $S_f = 0$.

@@

        The core of the semi-analytic modeling framework in \holodeck{} is implementing \eqref{eq:sam_final}.  There are classes for each of the key components: The galaxy stellar-mass function (GSMF, $\Phi$), the galaxy pair fraction ($F$), the merger timescale ($\tau$), and additional a scaling relation for mapping host galaxies via to MBH properties --- typically via an \mmbulge{} relation.  Each component class is implemented generically via basic API methods (e.g.~a GSMF that returns the number density of galaxies at a given stellar-mass and redshift), and can be subclassed to provide a particular implementation (e.g.~an analytic Schechter-function GSMF).  The number (or number density) of binaries is specified over a grid of parameters, typically in a four dimensional space of total mass ($M$), mass ratio ($q$), redshift ($z$), and binary rest-frame orbital frequency ($f_r$).

        The semi-analytic model formalism defines a continuous distribution of MBH binaries.  For many calculations, the discrete nature of MBHB sources is important, not only for examining individual continuous wave sources but also because of the important effects of discrete sources on the shape of the GWB due to the non-linear scaling of GW strain \citep{Sesana+2008}.  Additional, a component of the uncertainty in expected GW signals is due to `cosmic variance', which can be treated as Poisson noise when creating multiple samples (or `realizations') of our binary populations.  To handle discretization and resampling we use the \kalepy{} package \citep{kalepy2021}, in particular, a method of outlier sampling to capture regions of parameter space where finite number effects are important.  The \holodeck{} semi-analytic models specify the number (or number-density) of binaries over a grid, i.e.~$N_{ijkl}$.  By definition, grid cells which have an expected total number of binaries $\gg 1$ will not produce appreciable finite-number effects, so instead they are given a weighting based on a Poisson distribution centered around their expectation value, i.e.~$\mathcal{P}(N_{ijkl})$.  Grid cells with expected numbers of order unity or less do produce important finite-number effects, and thus are directly resampled

@@

    This research made use of \astropy, a community-developed core Python package for Astronomy \citep{astropy2013}, in addition to \scipy~\citep{scipy}, \ipython~\citep{ipython}, \jupyter~notebook~\citep{jupyter}, \numpy~\citep{numpy2011} \& \sympy~\citep{sympy2017}.  All figures were generated using \matplotlib~\citep{matplotlib2007}.  Kernel density estimation was performed using the \kalepy{} package (\href{https://github.com/lzkelley/kalepy}{github.com/lzkelley/kalepy}) \citep{kalepy2020}.

@@

>>>>

@@

@@

        Let $N(z)$ denote the number of binaries out to a redshift $z$, corresponding to a comoving volume $V_c$.
        We can calculate the characteristic strain from an ensemble of GW sources as \citep[][Eqs.~5/8]{Phinney-2001},
        	\begin{align}
        	h_c^2(f) = \int_0^\infty \!\! dz \; \frac{d^2 N}{dz \, d\ln f_r} \; h_s^2\lr{f_r}.
        	\end{align}
        In practice, the total number of sources in the Universe is calculated from the number-density, $n \equiv dN / dV_c$.  These can be related as \citep[][Eq.~6]{Sesana+2008},
        	\begin{align}
        	\frac{d^2 N}{dz \, d\ln f_r} = \frac{d n_c}{dz} \frac{dz}{dt} \frac{dt}{d\ln f_r} \frac{d V_c}{dz}.
        	\end{align}
        Note that the total number of binaries in a given frequency interval depends not only on the number density, but also their rate of frequency-evolution.  The faster binaries evolve through a given frequency interval, the fewer binaries will occupy that interval on average.  This is typically described by the binary `hardening' timescale (with respect to frequency), given by the expression,
        \begin{equation}
            \thardf \equiv \diffp{t}{{\ln\!f_r}} = \frac{f_r}{\partial f_r / \partial t}.
        \end{equation}
        Plugging in the relevant cosmographic relationships (see \secref{sec:app_eqs}) we have,
        	\begin{align}
        	h_c^2(f) & = \int_0^\infty \!\! dz \; \frac{dn_c}{dz} \, h^2\lr{f_r} \, 4\pi c \, d_c^2 \lr{1+z} \, \thardf.
        	\end{align}
        The challenge in modeling astrophysical sources of GWs is in calculating their number density.  Numerous approaches are possible, described in the following sections.

@@

        Define the distribution function of sources as $f(M,a,z) = d^2 n(M,a,z) / dM da$.  Here $M$ is the total mass of each systems, $a$ is the binary separation, and $z$ is the redshift.  Note that binary frequency $f_r$ could be used almost interchangeably with $a$, and that both evolve as a function of time or equivalently redshift.  This formalism can easily be extended to include other parameters, such as mass-ratio and eccentricity ($q$ and $e$).  We can write the conservation equation for binaries as of function of redshift as,
        \begin{equation}
            \label{eq:conservation}
            \diffp{f}{z} + \diffp{}{m} \lrs{f \diffp{m}{z}} + \diffp{}{a} \lrs{f \diffp{a}{z}} = S_f(m, a, z).
        \end{equation}
        Here $S_f$ is a source/sink function that can account for the creation or destruction of binaries.  Here we assume that all binary `formation' is encapsulated from binaries moving from one part of parameter space (i.e.~large separations and redshifts) to other parts of parameter space (i.e.~smaller separations and redshifts).  Thus we set $S_f = 0$.

@@

        The core of the semi-analytic modeling framework in \holodeck{} is implementing \eqref{eq:sam_final}.  There are classes for each of the key components: The galaxy stellar-mass function (GSMF, $\Phi$), the galaxy pair fraction \(F\), the merger timescale ($\tau$), and additional a scaling relation for mapping host galaxies via to MBH properties --- typically via an \mmbulge{} relation.  Each component class is implemented generically via basic API methods (e.g.~a GSMF that returns the number density of galaxies at a given stellar-mass and redshift), and can be subclassed to provide a particular implementation (e.g.~an analytic Schechter-function GSMF).  The number (or number density) of binaries is specified over a grid of parameters, typically in a four dimensional space of total mass ($M$), mass ratio ($q$), redshift ($z$), and binary rest-frame orbital frequency ($f_r$).

        The semi-analytic model formalism defines a continuous distribution of MBH binaries.  For many calculations, the discrete nature of MBHB sources is important, not only for examining individual continuous wave sources but also because of the important effects of discrete sources on the shape of the GWB due to the non-linear scaling of GW strain \citep{Sesana+2008}.  Additional, a component of the uncertainty in expected GW signals is due to `cosmic variance', which can be treated as Poisson noise when creating multiple samples (or `realizations') of our binary populations.  To handle discretization and resampling we use the \texttt{kalepy} package \citep{kalepy2021}, in particular, a method of outlier sampling to capture regions of parameter space where finite number effects are important.  The \holodeck{} semi-analytic models specify the number (or number-density) of binaries over a grid, i.e.~$N_{ijkl}$.  By definition, grid cells which have an expected total number of binaries $\gg 1$ will not produce appreciable finite-number effects, so instead they are given a weighting based on a Poisson distribution centered around their expectation value, i.e.~$\mathcal{P}(N_{ijkl})$.  Grid cells with expected numbers of order unity or less do produce important finite-number effects, and thus are directly resampled

@@

    This research made use of \texttt{astropy}, a community-developed core Python package for Astronomy \citep{astropy2013}, in addition to \texttt{Scipy}~\citep{scipy}, \texttt{ipython}~\citep{ipython}, \texttt{jupyter}~notebook~\citep{jupyter}, \texttt{Numpy}~\citep{numpy2011} \& \texttt{SymPy}~\citep{sympy2017}.  All figures were generated using \texttt{matplotlib}~\citep{matplotlib2007}.  Kernel density estimation was performed using the \texttt{kalepy}{} package (\href{https://github.com/lzkelley/kalepy}{github.com/lzkelley/kalepy}) \citep{kalepy2020}.

@@

    \section{Additional Equations}
        \label{sec:app_eqs}
        A binary is described by its total mass $M$, mass ratio $q \equiv m_2 / m_1$ (s.t.~$q\leq1$), (proper) semi-major axis $a$, and rest-frame orbital frequency $f_r$.  The characteristic gravitational wave mass is the chirp mass, $\mchirp \equiv M \, q^{3/5} \, \lr[-6/5]{1+q}$.

        The GW strain from a circular binary is \citep[][Eq.~7; sky and polarization averaged]{Sesana+2008},
            \begin{equation}
            \label{eq:source_strain}
            h_s = \frac{8}{10^{1/2}} \frac{\left(G\mchirp\right)^{5/3}}{c^4 \, d_L}
                \left(2 \pi f_r \right)^{2/3},
            \end{equation}
        for a luminosity distance $d_L$.  Recall that the luminosity distance and comoving distance are related as $d_L = d_c \, (1+z)$, that the comoving volume of the Universe is given by \citep{Hogg-1999},
        \begin{equation}
            \frac{d V_c}{dz} = 4\pi \frac{c}{H_0} \frac{d_c^2}{E(z)},
        \end{equation}
        and that redshift can be related to time (or age of the universe) as,
        \begin{align}
            \frac{dz}{dt} = H_0 \lr{1+z} E(z).
        \end{align}
        The evolution of the Hubble parameter over redshift can be calculated from the matter-energy constituents of the universe as,
        \begin{align}
            H(z) & = H_0(z) E(z), \\
            E(z) & \equiv \lrs[1/2]{\Omega_R \lr[4]{1+z} + \Omega_m \lr[3]{1+z} + \Omega_k \lr[2]{1+z} + \Omega_\Lambda}.
        \end{align}
        The $\Omega$ terms are the usual redshift-zero fractions of the critical density for radiation, non-relativistic matter, curvature, and dark energy.  We adopt the results of WMAP9 \citep{Hinshaw+2013} for our fiducial cosmology, assuming zero curvature: $H_0 = 69.33 \textrm{ km/s/Mpc}$, $\Omega_m = 0.2880$, $\Omega_\Lambda = 0.7120$.
